PineAPPL: combining EW and QCD corrections\\ for fast evaluation of LHC processes

### 2020-08-28 paper/sec-results.tex
ERN's final, very few, very minor typos

The values of the relevant physical parameters are chosen as
\begin{equation}
\begin{aligned}
M_\mathrm{W} &= \SI{80.419}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} &= \SI{91.176}{\giga\electronvolt} \text{,} \quad &
m_\mathrm{t} &= \SI{172.5}{\giga\electronvolt} \text{,} \quad \\
\Gamma_\mathrm{W} &= \SI{2.09291}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} &= \SI{2.49877}{\giga\electronvolt} \text{,} &
G_\mu &= \SI{1.16639e-5}{\per\giga\electronvolt\squared} \text{,}
\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{Z}$, $M_\mathrm{W}$, $m_\mathrm{t}$ are the values of the Z-boson, W-boson, and
top-quark masses, respectively, $\Gamma_\mathrm{Z}$ and $\Gamma_\mathrm{W}$ are the widths of
the Z and W bosons, and $G_\mu$ is value of the Fermi coupling. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.

The definition of observables and cuts is process-specific, and it follows the corresponding experimental measurements,
see section~\ref{subsec:processes_and_measurements}.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematic observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination would have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections, in particular
    about the fragmentation of partons into photons and vice-versa,
see refs.~\cite{Glover:1993xc,Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order of the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of \enquote{standard} EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases).

@@

%-------------------------------------------------------------------------------
\begin{figure}[!t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLASZHIGHMASS49FB}\\
    \caption{Validation and test of the \textsc{PineAPPL} grid for the ATLAS
      high-mass Drell--Yan lepton pair measurement in the high-mass region at
      a centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Aad:2013iua}. The first panel
      displays the relative difference (in per mille) between the {\sc PineAPPL}
      and the MC results for the central, upper and lower scale choices,
      for theories accurate to both NLO QCD and NLO QCD+EW. The second, third
      and fourth panels present the theoretical predictions, accurate to either
      NLO QCD or NLO QCD+EW, always normalised to the former; on top of the
      theoretical predictions, the PDF uncertainty, the scale uncertainty and
      the Monte Carlo uncertainty are displayed in turn. The relative
      uncertainty of the experimental data is also shown for comparison.}
    \label{fig:atlaszhighmass49fb}
\end{figure}
%-------------------------------------------------------------------------------

@@

The NLO QCD+EW corrections always lead to a reduction of the cross section in
comparison to the NLO QCD prediction. The size of this shift is comparable to
the data uncertainty at small values of $M_{\ell\bar\ell}$, and rapidly becomes
negligible with respect to it as the value of the invariant mass increases
and the data uncertainty blows up. This fact suggests a couple of observations
in light of the inclusion of EW corrections in a fit of PDFs. First, the
description of the more precise bins in the low invariant mass range is likely
to change, and will possibly become more accurate should the inclusion of EW
corrections improve the data/theory agreement. Second, the kinematic cut that
excludes any data point at large $M_{\ell\bar\ell}$ can be safely removed: any
shift in the predictions induced by the more accurate NLO QCD+EW theory is
likely to be easily accommodated by the large data uncertainty. However, 
due to the increased Run-II LHC luminosity data will become more precise.

>>>>

The values of the relevant physical parameters are chosen as
\begin{equation}
\begin{aligned}
M_\mathrm{W} &= \SI{80.419}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} &= \SI{91.176}{\giga\electronvolt} \text{,} \quad &
m_\mathrm{t} &= \SI{172.5}{\giga\electronvolt} \text{,} \quad \\
\Gamma_\mathrm{W} &= \SI{2.09291}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} &= \SI{2.49877}{\giga\electronvolt} \text{,} &
G_\mu &= \SI{1.16639e-5}{\per\giga\electronvolt\squared} \text{,}
\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{Z}$, $M_\mathrm{W}$, $m_\mathrm{t}$ are the values of the Z-boson, W-boson, and
top-quark masses, respectively, $\Gamma_\mathrm{Z}$ and $\Gamma_\mathrm{W}$ are the widths of
the Z and W bosons, and $G_\mu$ is the value of the Fermi coupling. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.

The definition of observables and cuts is process-specific, and it follows the corresponding experimental measurements,
see section~\ref{subsec:processes_and_measurements}.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematic observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination would have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections, in particular
    about the fragmentation of partons into photons and vice-versa,
see refs.~\cite{Glover:1993xc,Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order of the EW corrections, they are not included in our
computations. In fact, while nothing prevents one to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of \enquote{standard} EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases).

@@

%-------------------------------------------------------------------------------
\begin{figure}[!t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLASZHIGHMASS49FB}\\
    \caption{Validation and test of the \textsc{PineAPPL} grid for the ATLAS
      Drell--Yan lepton pair measurement in the high-mass region at
      a centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Aad:2013iua}. The first panel
      displays the relative difference (in per mille) between the {\sc PineAPPL}
      and the MC results for the central, upper and lower scale choices,
      for theories accurate to both NLO QCD and NLO QCD+EW. The second, third
      and fourth panels present the theoretical predictions, accurate to either
      NLO QCD or NLO QCD+EW, always normalised to the former; on top of the
      theoretical predictions, the PDF uncertainty, the scale uncertainty and
      the Monte Carlo uncertainty are displayed in turn. The relative
      uncertainty of the experimental data is also shown for comparison.}
    \label{fig:atlaszhighmass49fb}
\end{figure}
%-------------------------------------------------------------------------------

@@

The NLO QCD+EW corrections always lead to a reduction of the cross section in
comparison to the NLO QCD prediction. The size of this shift is comparable to
the data uncertainty at small values of $M_{\ell\bar\ell}$, and rapidly becomes
negligible with respect to it as the value of the invariant mass increases
and the data uncertainty blows up. This fact suggests a couple of observations
in light of the inclusion of EW corrections in a fit of PDFs. First, the
description of the more precise bins in the low invariant mass range is likely
to change, and will possibly become more accurate should the inclusion of EW
corrections improve the data/theory agreement. Second, the kinematic cut that
excludes any data point at large $M_{\ell\bar\ell}$ can be safely removed: any
shift in the predictions induced by the more accurate NLO QCD+EW theory is
likely to be easily accommodated by the large data uncertainty. However, 
due to the increased Run-II LHC luminosity, data will become more precise.
